\chapter{Checklist}
\label{chap:appendix:checklist}
Use the following list to check if you have followed the hints from this document in your work.
Refer to \cref{chap:introduction} for more information on the items.

\tabulinesep=2.5mm
% tabu allows to use relative width specifier: Use X[<ratio>] to specify the width of a column. In
% this example, the table is divided into 20 parts that are spread over the three columns.
\begin{longtabu} to 0.8\textwidth {rX}
% Begin header on first page
\caption[Formatting checklist]{The checklist for correctly formatted and prettier documents.}
\label{tab:checklist}
\\ \addlinespace
\endfirsthead
% End header on first page

% Begin header on consecutive pages
\caption[]{The checklist for correctly formatted and prettier documents, continued.}
\\ \addlinespace
\toprule
\endhead
% End header on consecutive pages

% Begin footer
\\ \addlinespace
\multicolumn{2}{c}{Table is continued on the next page.} \\
\endfoot
% End footer

% Begin footer on last page
\endlastfoot
% End footer on last page

% Begin content
\toprule
1. & Check for incorrect/missing citations (\emph{(?)}) or references (\emph{??}).\\
\midrule
2. & Remove all \LaTeX{} errors and underfull/overfull boxes.\\
\midrule
3. & Make sure to use a consistent encoding for your \TeX{} files, especially for special characters (ä, ö, ü, ß, \dots).\\
\midrule
4. & Format your \textsc{Bib}\TeX{} document: Check if the information is correct and complete for each entry. Pay notice to warnings when running  the \texttt{bibtex} command.\\
\midrule
5. & Add access dates to online sources in your bibliography (see \texttt{library.bib}) or in footnotes.\\
\midrule
6. & Run a spell checker over your document (included with some \LaTeX{} editors).\\
\midrule
7. & Place figures or tables in the \TeX{} file at the end of the paragraph you are referring to them in the text (\texttt{\textbackslash{}cref}).\\
\midrule
8. & Use the correct format for quotation marks (\emph{``x''} or \emph{\glqq{}x\grqq{}}).\\
\midrule
9. & Use the correct format for separating paragraphs (one empty line). Manual line breaks (\texttt{\textbackslash{}\textbackslash}) should be avoided.\\
\midrule
10. & Use the correct form of dashes (-, --, ---).\\
\midrule
11. & Name sources of images/figures you did not create yourself in the description (\emph{Source: [X]}).\\
\midrule
12. & Use the short form of captions for List of Figures, Tables, etc. (\texttt{caption[<short>]\{<long>\}}).\\
\midrule
13. & If you print your work double-sided (recommended for bachelor and master thesis) remove the \texttt{oneside} option from the document class.\\
\midrule
14. & Make sure to use high-quality figures and images that are readable both in the digital and the printed version.\\
\midrule
15. & If you include the declaration of honor do not forget to sign it.\\
\bottomrule
% End content
\end{longtabu}